Optimal strategies in mental rotation tasks

### 2013-05-13 man/final-project.tex
Add figures
Remove random comments
Add abstract with TODO
Fix typo
Clean up computational model section

\usepackage{amsmath, amsthm, amssymb}
% \usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}

@@

% \usepackage{etoolbox}
% \makeatletter
% \patchcmd{\thebibliography}{%
%   \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
%   \subsubsection*{References}}{}{}
% \makeatother

@@

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}

@@

% \begin{wrapfigure}{r}{0.4\textwidth}
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
  \end{center}
  \caption{Classic mental rotation task from \cite{Shepard1971}.}
  \label{fig:mental-rotation}
% \end{wrapfigure}
\end{figure}

@@

qIn the case of the classic mental rotation task, we might ask: in what
direction should the object be rotated?  What are the requirements for
``congruency''? When should one stop rotating and accept the
hypothesis that the objects are different? In this project, I will
investigate the optimal computational solution to this task. Thus, the
specific question this project aims to answer is: what is the rational
computational solution to the mental rotation task, and does it
predict the people's behavior on the task?

@@

\subsection{Mental rotation}

\subsection{Optimal experiment design}
% We do not know the form of $S$, and so cannot compute this integral
% analytically. We cannot rely on a simple Monte Carlo simulation to
% estimate it, either: because mental rotation is costly in terms of
% cognitive resources and must be performed in a sequence, we cannot
% evaluate $S(I_b|I_M)$ at arbitrary $I_M$. Instead, we must choose a
% relatively small, sequential set of $I_M$ to estimate the
% integral.

@@

@@

@@

\subsection{Likelihood when the configurations are distinct}

The likelihood is easy to compute under $h_0$, because we assume that
$X_a$ and $X_b$ are independent. Thus:
\begin{equation}
  p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)
\end{equation}

@@

However, we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
introduce a new variable $X_R$ denoting a mental image, which
approximates $RX_a$. The $X_R$ are generated sequentially by repeated
application of a function $\tau$:
\begin{align*}
  X_R&=RX_a\\
  &=\tau(X_{R-r}, r)\\
  &=\tau(\tau(X_{R-2r}, r), r)\\
  &\ldots{}\\
  &=\tau^{(\frac{R}{r})}(X_a, r)
\end{align*} 
Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
applications of $\tau$. Using this sequential function, we get:
\begin{align}
  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  x&= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
\end{align}

\subsection{Likelihood ratio}

Once we have computed both likelihoods, we compute their ratio:
\begin{equation}
  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}
\end{equation} 
If $\ell<1$, then $h_0$ is the more likely hypothesis. If $\ell>1$,
then $h_1$ is the more likely hypothesis.

@@

\subsection{Likelihood when the configurations are distinct}

We define the prior probabilities over stimuli based on how they are
generated. For shapes with $n$ vertices, each vertex is at a random
angle with a radius random chosen between 0 and 1, in polar
coordinates, and could be chosen in any of $n!$ different ways. Thus,
the prior over shapes is:
\begin{equation}
  p(X)=n!\left(\frac{1}{2\pi}\right)^n
\end{equation} 
And so the joint likelihood is just:
\begin{equation}
  p(X_a, X_b\ \vert \ h_0)=\left[n!\left(\frac{1}{2\pi}\right)^n\right]^2
\end{equation}

@@

@@

@@

@@

\section{Method}

% \renewcommand\bibsection{\subsubsection*{\refname}}
\renewcommand\refname{\normalsize{References}}
\bibliographystyle{ieeetr}
\bibliography{references}
% - References. Give at least three references cited in describing the
% background to your project.

>>>>

\usepackage{amsmath, amsthm, amssymb}
% \usepackage{natbib}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{subcaption}

@@

@@

\newcommand{\TODO}[1]{\textcolor{red}{[TODO: #1]}}

@@

\begin{abstract}
\TODO{}
\end{abstract}

@@

In the case of the classic mental rotation task, we might ask: in what
direction should the object be rotated?  What are the requirements for
``congruency''? When should one stop rotating and accept the
hypothesis that the objects are different? In this project, I will
investigate the optimal computational solution to this task. Thus, the
specific question this project aims to answer is: what is the rational
computational solution to the mental rotation task, and does it
predict the people's behavior on the task?

@@

% We do not know the form of $S$, and so cannot compute this integral
% analytically. We cannot rely on a simple Monte Carlo simulation to
% estimate it, either: because mental rotation is costly in terms of
% cognitive resources and must be performed in a sequence, we cannot
% evaluate $S(I_b|I_M)$ at arbitrary $I_M$. Instead, we must choose a
% relatively small, sequential set of $I_M$ to estimate the
% integral.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=0.95\textwidth]{../figures/shepard-rotation.png}
  \caption{Classic mental rotation task from \cite{Shepard1971}.}
  \label{fig:mental-rotation}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/stimuli_shapes.pdf}
    \vspace{0pt}
    \caption{Example stimuli}
    \label{fig:stimuli}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/likelihood_function.pdf}
    \caption{Likelihood function}
    \label{fig:likelihood}
  \end{subfigure}
  \caption{Rotated shapes and corresponding similarities.}
\end{figure}

@@

The likelihood is easy to compute under $h_0$, because we assume that
$X_a$ and $X_b$ are independent. Thus:
\begin{equation}
  p(X_a,\ X_b\ \vert \ h_0)=p(X_a)p(X_b)
  \label{eq:lh-h0}
\end{equation}

@@

However, we cannot compute $p(X_b\vert X_a, R)$ directly. Instead, we
introduce a new variable $X_R$ denoting a mental image, which
approximates $RX_a$. The $X_R$ are generated sequentially by repeated
application of a function $\tau$:
\begin{align}
  X_R&=RX_a\nonumber \\
  &=\tau(X_{R-r}, r)\nonumber \\
  &=\tau(\tau(X_{R-2r}, r), r)\nonumber \\
  &\ldots{}\nonumber \\
  &=\tau^{(\frac{R}{r})}(X_a, r)
  \label{eq:tau}
\end{align} 
Where $r$ is a small angle, and $\tau^{(i)}$ indicates $i$ recursive
applications of $\tau$. Using this sequential function, we get:
\begin{align}
  p(X_a, X_b\ \vert \ h_1)&=\int_R \int_{X} p(X_b\vert X) p(X\vert X_a, R)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  x&= \int_R \int_X p(X_b\vert X)\delta(\tau^{(\frac{R}{r})}(X_a, r)-X)p(X_a)p(R)\ \mathrm{d}X\ \mathrm{d}R \nonumber \\
  &= \int_R p(X_b\vert X_R)p(X_a)p(R)\ \mathrm{d}R
  \label{eq:lh-h1}
\end{align}

Once we have computed both likelihoods, we compute their ratio:
\begin{equation}
  \ell=\frac{p(X_a, X_b\ \vert \ h_1)}{p(X_a, X_b\ \vert \ h_0)}=\frac{\int_R p(X_b\ \vert\ X_R)p(R)\ \mathrm{d}R}{p(X_b)}
  \label{eq:lh-ratio}
\end{equation}
If $\ell<1$, then $h_0$ is the more likely hypothesis. If $\ell>1$,
then $h_1$ is the more likely hypothesis.

@@

We define the prior probabilities over stimuli based on how they are
generated. For shapes with $n$ vertices, each vertex is at a random
angle with a radius random chosen between 0 and 1, in polar
coordinates, and could be chosen in any of $n!$ different ways. Thus,
the prior over shapes is:
\begin{equation}
\subsection{Likelihood when the configurations are distinct}

  p(X)=n!\left(\frac{1}{2\pi}\right)^n
\end{equation} 
And so the joint likelihood is just:
\begin{equation}
  p(X_a, X_b\ \vert \ h_0)=\left[n!\left(\frac{1}{2\pi}\right)^n\right]^2
\end{equation}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/li_regression.pdf}
    \caption{Linear interpolation}
    \label{fig:li}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{../figures/vm_regression.pdf}
    \caption{Parametric regression}
    \label{fig:vm}
  \end{subfigure}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{../figures/bq_regression.pdf}
  \caption{Bayesian Quadrature}
  \label{fig:bq}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{../figures/model_rotations.pdf}
  \caption{Model rotations}
  \label{fig:rotations}
\end{figure}

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{../figures/Z_accuracy.pdf}
  \caption{Accuracy in estimating $Z$.}
  \label{fig:accuracy}
\end{figure}

@@

% \renewcommand\bibsection{\subsubsection*{\refname}}
\renewcommand\refname{\normalsize{References}}
\bibliographystyle{ieeetr}
\bibliography{references}
